---
title: "P8105 Homework 2"
author: "Maya Krishnamoorthy"
date: "2024-09-25"
output: github_document
---

``` {r setup, echo=FALSE, message=FALSE}
library(tidyverse)
library(readxl)
```

## Problem 1

This problem focuses on NYC transit data. The dataset `NYC_Transit_Subway_Entrance_and_Exit_Data.csv` contains information related to each entrance and exit for each subway station in NYC.

*Step 1*: Read and clean the data; retain line, station, name, station latitude / longitude, routes served, entry, vending, entrance type, and ADA compliance. Convert the entry variable from character (YES vs NO) to a logical variable (the ifelse or case_match function may be useful).

``` {r}
nyc_transit = 
  read_csv("data/NYC_Transit_Subway_Entrance_and_Exit_Data.csv") |> 
  janitor::clean_names() |> 
  select(line, station_name, station_latitude, station_longitude, route1:route11, entry, vending, entrance_type, ada) |> 
  mutate(
    entry = case_match(
      entry, 
      "YES" ~ TRUE, 
      "NO" ~ FALSE)
  )
```

*Step 2*: Write a short paragraph about this dataset – explain briefly what variables the dataset contains, describe your data cleaning steps so far, and give the dimension (rows x columns) of the resulting dataset. Are these data tidy?

